Diffusive Homeostasis in a Self-Organizing Recurrent Neural Network \\ \begin{large} Spatially dependent Interaction as a Determinant of Neural Activity and Plasticity \end{large}

### 2017-03-09 draft.tex
.

\subsubsection{Former Single-Cell Homeostasis as an Integral Control}
Former versions of the LIF-SORN and SORN used a simple form of integral control to regulate excitatory firing rates. Individual firing rates of excitatory cells were estimated by a spike counter over short subsequent time periods and compared to a target value. This difference was then fed to a controller which increased or decreased neuronal excitability in an additive way (making it a discrete form of integral control). As mentioned before, this type of control allowed to reliably tune each cell to a given target firing rate under changing input. A global target firing rate was chosen for all excitatory neurons. Consequently, this resulted in a very narrow distribution of excitatory firing rates.

@@

In the plastic network, the absence of this breaking of spatial homogeneity even led to an under-representation of bidirectional connections, which is known to be an effect of STDP in recurrent networks \cite{Syn_Plast_Abbott}. The separation between simulations with and without a spatial connection profile is retained after having implemented diffusive homeostasis, as shown in Figure \ref{RC_CFB_plot}. 
\begin{figure}
%%% plot_RC_CFB.py
\includegraphics[width=\textwidth]{../../plots/syn_topology/RC_CFB_plot.png}
\caption{Fraction of bidirectional connections normalized over the expected fraction for a random graph with equal total connection fraction. Data from 10 trials, linewidth represents standard error.}
\label{RC_CFB_plot}
\end{figure}
\subsubsection{Distribution of Synaptic Weights} \label{Syn_Weight_Dist_Section}
A well-studied property of neural networks is the distribution of synaptic weights. Experimental studies in hippocampus and cortical regions suggest heavy-tailed log-normal-like distributions of excitatory synaptic efficacies \cite{Song_Connectivity_2005,Lisman_Synapses_1993,Yasumatsu_Synapses_2008,Loewenstein_Spine_Sizes}. Theoretical models are mainly based on a combination of multiplicative and additive weight dynamics \cite{Loewenstein_Spine_Sizes,Statman_Synapses_2014}, which is in line with our implementation of multiplicative normalization and additive STDP. Figure \ref{Weight_Dist} compares the resulting distributions. We found that we can retain a log-normal-like distribution in combination with diffusive homeostasis, giving a similar distribution as for non-diffusive homeostasis.
\begin{figure}
%%% plot_weight_dist_mult_sets.py
\includegraphics[width=\textwidth]{../../plots/syn_topology/syn_weight_dist_mult_sets.png}
\caption{Distribution of decadic logarithm of excitatory synaptic weights at $\mathrm{t=1500\,s}$ for diffusive and non-diffusive homeostasis (both with distance-dependent connectivity). Gaussian fits returned $\mathrm{\mu_{\rm diff.}=-0.128\, \mathrm{log}_{10}(mV)}$, $\mathrm{\mu_{\rm non-diff.}=-0.064\, \mathrm{log}_{10}(mV)}$, $\mathrm{\sigma_{\rm diff.}=0.415\, log_{10}(mV)}$ and $\mathrm{\sigma_{non-diff.}=0.395\, log_{10}(mV)}$. Weight matrices were taken from 5 simulations, each at $\rm t = 1500\,s$.}
\label{Weight_Dist}
\end{figure}

@@

We made use of the \textit{powerlaw python package} for this purpose \cite{Powerlaw_Package,Alstott_2014}. It provides a toolset for the analysis of potentially power-law distributed data, applying the maximum likelihood method as described by Clauset et al. \cite{Clauset_2009}. An additional feature of this package---also originally proposed by Clauset et al.---is the possibility to automatically find a lower bound $x_{\rm min.}$ below which data points are discarded from the analysis, based on the goodness of the estimated power law to the actual data. The authors argue that this feature accounts for the possibility that smaller values of a dataset might not as strictly follow a power law as found in the tail of the distribution. This optimization is based on the Kolmogorov-Smirnov statistic, which is the maximum absolute distance $d_{\rm max}$ between the empirical cumulative distribution function of the data and the cumulative distribution function of the analytic prediction. It can be further used to perform a statistical test, allowing a judgment whether a given dataset of random variables was generated from the given PDF \cite{EncyMath_Kolmogorov_Smirnov}. For the sake of finding the optimal value of $x_{\rm min.}$ however, the authors simply minimize $d_{\rm max}$ in order to find the best fitting setting.

Still, estimating an power-law exponent based on the given data required setting an upper and lower limit over which

This feature was preserved under diffusive homeostasis, as shown in Figure \ref{Syn_Lifetimes}. The resulting slopes are within previously reported values of $\mathrm{\approx - 5/3}$ \cite{SORN_Paper}.
\begin{figure}
%%% plot_syn_lifetimes.py
\includegraphics[width=\textwidth]{../../plots/syn_topology/syn_lifetimes_new.png}
\caption{Distribution of lifetimes of recurrent excitatory synaptic connections. Linear fits ignoring falloff (only using bins with $\rm t \leq 150 \, s$) for large synaptic lifetimes resulted in a power-law exponent of $\mathrm{\alpha_{\rm Diff.} = -1.455}$ (standard error $\mathrm{\sigma_{\rm Diff.} = 5.52 \cdot 10^{-4}}$) for diffusive homeostasis and $\mathrm{\alpha_{\rm Non-Diff.} = -1.468}$ (standard error $\mathrm{\sigma_{\rm Non-Diff.} = 5.70 \cdot 10^{-4}}$) for non-diffusive homeostasis.}
\label{Syn_Lifetimes}
\end{figure}
To get an approximative explanation for the observed slope in the region of small lifetimes, we note that short synaptic lifetimes are most likely coming from synapses whose weights did not reach large values compared to the initial weight. The effect of multiplicative normalization is small compared to additive STDP in this regime of small weights. Small weights are therefore determined by additive quasi-random (due to their weak coupling to the postsynaptic neuron) fluctuations, which, as a very simple approximation, can be described as a Brownian motion with a certain standard deviation $\mathrm{\sigma}$. The first-hitting time of this Brownian motion is described by a Lévy distribution \cite{Roy_1968}:
\begin{equation}
p(t,w_0,w_{\rm prune}) = \frac{ | w_0-w_{\rm prune}|}{\sqrt{2\pi \sigma t^3}} \mathrm{exp} \left( \frac{ \left( w_0-w_{\rm prune} \right) ^2}{2\sigma t}\right)
\label{Levy_Dist}
\end{equation}
The term $\mathrm{t^{-3/2}}$ dominates for larger $\mathrm{t}$, which very roughly resembles the power-law exponents found in Figure \ref{Syn_Lifetimes}. Zheng et al. found a similar slope in the binary version of the SORN and also argued that this could be referred to random-walk behavior \cite{Pengsheng_2013}. Still, understanding the exponent of this distribution is subject to ongoing research and it has been argued that a more precise explanation could be used to relate certain parameters of synaptic plasticity to the value of this exponent. Especially, the balance between long-term potentiation and depression is thought to influence this exponent, where potentiation/depression dominated plasticity leads to a smaller/larger absolute value of the exponent of the distribution.

@@

We noted further that the difference between shuffled matrices corresponding to non-diffusive and diffusive homeostasis is in line with our observation from Figure \ref{Weight_Dist}, which indicated a broader distribution of weights for diffusive homeostasis.
\begin{figure}
%%% plot_sum_outgoing_weights_dist.py
\includegraphics[width=\textwidth]{../../plots/syn_topology/outgoing_weights_distribution.png}
\caption{Mean of outgoing weights of excitatory neurons in order sorted by magnitude. Dashed lines are means obtained by randomly shuffling synaptic weights. Data was taken from a single weight matrix each, at $\mathrm{t = 1500\,s}$.}
\label{Out_Weight_Mean_Quantile}
\end{figure}

@@

Figure \ref{dynamics_rate_threshold_dyn_mean_field_sweeney} depicts the resulting dynamics of excitatory rates and thresholds for $\mathrm{\alpha=\lbrace 0.5,0.9,1.0\rbrace}$. We ran simulations with other values of $\rm \alpha$ and found that all rates approach the same target rate of $\mathrm{3\,Hz}$ and thresholds move towards the same configuration for either choice of $\mathrm{\alpha}$ except for the limiting case of $\mathrm{\alpha=1}$. A significant difference only exists within the transient dynamics leading to the steady state. Roughly speaking, a smaller value of $\mathrm{\alpha}$ leads to a faster relaxation.
\begin{figure}
%%% plot_sim_mean_field_sweeney.py
\includegraphics[width=\textwidth]{../../plots/rate_threshold_simpl_hip_alpha/rate_threshold_simpl_hip_alpha_comb}
\caption{Dynamics of rates (A) and thresholds (B) of excitatory population of 400 Neurons and different values of $\mathrm{\alpha}$ (see Equations \eqref{dyn_rate_equation} and \eqref{diff_hom_simpl_sweeney}).}
\label{dynamics_rate_threshold_dyn_mean_field_sweeney}
\end{figure}
These findings led us to the conclusion that---apart from the special case $\mathrm{\alpha=1}$, which is the equivalent to instantaneous diffusion---a mean-field model describing the diffusive signal as a mixture of individual activity and all neurons is not suitable for explaining the existence of stable heterogeneous distributions of rates for a broad range of diffusion constants. We did not intent to debase the validity of results of the mean-field model of Sweeney et al. However, we could show that the steady-state solution of this simplified model will result in the same firing rate $\phi_0$ for all neurons, given any value of $\mathrm{\alpha}$ except $\mathrm{\alpha=1}$. This can be seen by setting the left hand side of Equation \eqref{diff_hom_simpl_sweeney} to $0$ (which is necessarily the case in the steady state) and rearranging the equation:
\begin{equation}
(\alpha-1)\frac{\phi_i - \phi_0}{\phi_i} = \alpha \frac{\langle \phi \rangle - \phi_0}{\langle \phi \rangle}
\label{diff_hom_simpl_sweeney_2}
\end{equation}
The right term of the equation is the same for all neurons. Since the left term is monotonically increasing as a function of $\phi_i$ as long as $\mathrm{\alpha<1}$, only one specific solution $\phi_i = \Phi$ for all $i$ exists that equals the given term on the right. This implies $\langle \phi \rangle = \Phi$. Thus,
\begin{equation}
(\alpha-1)\frac{\Phi - \phi_0}{\Phi} = \alpha \frac{\Phi - \phi_0}{\Phi}
\label{diff_hom_simpl_sweeney_3}
\end{equation}
which is only fulfilled for $\Phi = \phi_0$.

@@

\section{Discussion}

@@

\subsection{The Firing Rate Distribution} \label{Fir_Rate_Dist_Discussion_Section}
By implementing diffusive homeostasis into the LIF-SORN model we managed to achieve a broader, skewed firing rate distribution while preserving appreciated features of network topology. Though the distribution did approximately resemble a log-normal-like shape, it did not perfectly match our expectations. In order to acquire a deeper understanding of the mechanisms that determine these statistics we derived an analytical expression that allowed us to reliably predict the steady state of firing rates based on the spatial structure of excitatory neurons. We would like to emphasize that this explanation differs from common approaches of understanding the emergence of heavy-tailed statistics of firing rates in recurrent networks. Theoretical studies on the behavior of recurrent networks are largely based on the assumption that the excitability of neurons is not individually fine-tuned but rather randomly distributed and adjusted globally to achieve a desired mean firing rate. Predictions about the resulting activity are then made based on (potentially simplified) statistical features of recurrent connectivity and the specific form of the neuronal transfer function used in the model, see e.g. \cite{Roxin_Firing_Rate_Distribution,Vreeswijk1998}. Our interpretation of our results leads to the conclusion that this canonical approach matches the limiting case of instant diffusion: While Figures \ref{Fir_Rate_Dist_Instant_compare} and \ref{Fir_Rate_Dist_Compare} did not reveal many differences in terms of overall statistics, the decorrelation discussed in Section \ref{Section_Rand_Mat_vs_Sim} and plotted in Figure \ref{Corr_Coeff_vs_D} suggests that increasing the diffusion constant goes along with a transition between a spatially determined configuration of excitatory firing rates and a network topology-determined behavior.

@@

\subsection{Network Topology} 
Even though our investigations on network topology were rather intended as a form of sanity check with respect to properties already found in earlier versions, we found some differences compared to non-diffusive homeostasis.

\subsubsection{Distribution of Weights}
In Section \ref{Syn_Weight_Dist_Section}, we presented our results regarding the distribution of weights, which turned out to be broader but still approximately log-normal distributed for diffusive homeostasis. As a first interpretation, we hypothesized that this might due to a slightly decreased amount of STDP-induced potentiation (Figure \ref{STDP_change}) in the case of diffusive homeostasis in the first half of the simulation. Another explanation one might consider is that the increased firing rate for some neurons in the case of diffusive homeostasis lead to a greater amount of fluctuation because of the increased amount of STDP-events. However, since the overall variance of STDP-induced fluctuations was not significantly higher but even slightly lower compared to non-diffusive homeostasis, this explanation seems rather unlikely.

Taking a closer look at the distributions depicted in Figure \ref{Weight_Dist} we noted that the increased width of the distribution is largely due to an increment on the left half of the curve. Since new weights are initialized at $\mathrm{10^{-4}\, mV}$ this might indicate that some weights are drifting much more slowly from their value of initialization compared to other weights. These ``slow drifters" are presumably connecting neurons both spiking at a below-average firing rate.
This provides an explanation why this effect is more present in the case of diffusive homeostasis since it causes a portion of neurons to have rather low firing rates. On the other hand, for non-diffusive homeostasis, one can assume that the amount of STDP-fluctuations is approximately equal for all recurrent excitatory weights because of the fixation of firing rates.

\subsubsection{Strong Differences in Outgoing Weights}
The most prominent difference between homeostatic mechanisms we found within network topology was the emergence of significantly above-average mean outgoing weights for a small subset of neurons. By comparing the given mean values to statistics acquired by randomly shuffling recurrent excitatory weights (see Figure \ref{Out_Weight_Mean_Quantile}), we found that this effect cannot be explained by the overall increase of weight heterogeneity that we discussed in the previous section. Thus, we identify the over-representation of strong outgoing weights as a non-random effect. However, we would like to emphasize that diffusive homeostasis is only indirectly responsible in the sense that it allows for a sustained heterogeneity among firing rates.

>>>>

\subsubsection{Single-Cell Homeostasis as an Integral Control}
Former versions of the LIF-SORN and SORN used a simple form of integral control to regulate excitatory firing rates. Individual firing rates of excitatory cells were estimated by a spike counter over short subsequent time periods and compared to a target value. This difference was then fed to a controller which increased or decreased neuronal excitability in an additive way (making it a discrete form of integral control). As mentioned before, this type of control allowed to reliably tune each cell to a given target firing rate under changing input. A global target firing rate was chosen for all excitatory neurons. Consequently, this resulted in a very narrow distribution of excitatory firing rates.

@@

In the plastic network, the absence of this breaking of spatial homogeneity even led to an under-representation of bidirectional connections, which is known to be an effect of STDP in recurrent networks \cite{Syn_Plast_Abbott}. The separation between simulations with and without a spatial connection profile is retained after having implemented diffusive homeostasis, as shown in Figure \ref{RC_CFB_plot}. 
\begin{figure}
%%% plot_RC_CFB.py
\includegraphics[width=\textwidth]{../../plots/syn_topology/RC_CFB_plot.png}
\caption{Fraction of bidirectional connections normalized over the expected fraction for a random graph with equal total connection fraction. Data from 10 trials, linewidth represents standard error.}
\label{RC_CFB_plot}
\end{figure}
\subsubsection{Distribution of Synaptic Weights} \label{Syn_Weight_Dist_Section}
A well-studied property of neural networks is the distribution of synaptic weights. Experimental studies in hippocampus and cortical regions suggest heavy-tailed log-normal-like distributions of excitatory synaptic efficacies \cite{Song_Connectivity_2005,Lisman_Synapses_1993,Yasumatsu_Synapses_2008,Loewenstein_Spine_Sizes}. Theoretical models are mainly based on a combination of multiplicative and additive weight dynamics \cite{Loewenstein_Spine_Sizes,Statman_Synapses_2014}, which is in line with our implementation of multiplicative normalization and additive STDP. Figure \ref{Weight_Dist} compares the resulting distributions. We found that we can retain a log-normal-like distribution in combination with diffusive homeostasis, giving a similar distribution as for non-diffusive homeostasis. Both distributions were slightly left-skewed ($\mathrm{v_{\rm diff.} = -1.21}$,$\mathrm{v_{\rm non-diff.} = -1.44}$). 
\begin{figure}
%%% plot_weight_dist_mult_sets.py
\includegraphics[width=\textwidth]{../../plots/syn_topology/syn_weight_dist_mult_sets.png}
\caption{Distribution of decadic logarithm of excitatory synaptic weights at $\mathrm{t=1500\,s}$ for diffusive and non-diffusive homeostasis (both with distance-dependent connectivity). Gaussian fits (non-linear least squares) returned $\mathrm{\mu_{\rm diff.}=-0.128\, \mathrm{log}_{10}(mV)}$, $\mathrm{\mu_{\rm non-diff.}=-0.064\, \mathrm{log}_{10}(mV)}$, $\mathrm{\sigma_{\rm diff.}=0.415\, log_{10}(mV)}$ and $\mathrm{\sigma_{non-diff.}=0.395\, log_{10}(mV)}$. Weight matrices were taken from 5 simulations, each at $\rm t = 1500\,s$.}
\label{Weight_Dist}
\end{figure}

@@

We made use of the \textit{powerlaw python package} for this purpose \cite{Powerlaw_Package,Alstott_2014}. It provides a toolset for the analysis of potentially power-law distributed data, applying the maximum likelihood method as described by Clauset et al. \cite{Clauset_2009}. An additional feature of this package---also originally proposed by Clauset et al.---is the possibility to automatically find a lower bound $x_{\rm min.}$ below which data points are discarded from the analysis, based on the goodness of the estimated power law to the actual data. The authors argue that this feature accounts for the possibility that smaller values of a dataset might not as strictly follow a power law as found in the tail of the distribution. This optimization is based on the Kolmogorov-Smirnov statistic, which is the maximum absolute distance $d_{\rm max.}$ between the empirical cumulative distribution function of the data and the cumulative distribution function of the analytic prediction. It can be further used to perform a statistical test, allowing a judgment whether a given dataset of random variables was generated from the given PDF \cite{EncyMath_Kolmogorov_Smirnov}. For the sake of finding the optimal value of $x_{\rm min.}$ however, the authors simply minimize $d_{\rm max.}$ in order to find the best fitting setting.

Observing the distribution of synaptic lifetimes, we found that fitting a power law to the simulation data would rather require to define an \emph{upper} bound $x_{\rm max.}$ than a lower one, since we observed a falloff from the linear behavior in a log-log plot for large synaptic lifetimes, see Figure \ref{Syn_Lifetimes}. The \textit{powerlaw package} allows manually setting an upper bound, but the optimization described before is only available for $x_{\rm min.}$. We therefore decided to manually implement the same approach for $x_{\rm max.}$, while keeping the lower bound fixed at $x_{\rm min.} = 1 \, s$.
We swept through values for $x_{\rm max.}$ within $\mathrm{2\,s} \leq x_{\rm max.} \leq \mathrm{750\,s}$, with $\rm 750 \, s$ being the maximal possible synaptic lifetime to be recorded. We estimated the power-law exponent with the aforementioned package for each value of $x_{\rm max.}$ and calculated $d_{\rm max.}$. The value of $x_{\rm max.}$ resulting in the smallest $d_{\rm max.}^{min.}$ was then picked for the final estimation of $\alpha$. We did expect $d_{\rm max.}$ to go through a minimum for a certain value of $x_{\rm max.}$ and to increase again due to the statistical deviations for long lifetimes. We did not observe this behavior, however. Instead, $d_{\rm max.}$ was monotonically decreasing but converging. Therefore, we actually found its minimal value for $x_{\rm max.} = \rm 750 \, s$ for both diffusive homeostasis and non-diffusive homeostasis ($d_{\rm max.,diff.}^{min.} = 0.30$,$d_{\rm max.,non-diff.}^{min.} = 0.34$). We thus accepted the values for $\alpha$ acquired by MLE for both entire datasets as valid, yielding $\mathrm{\alpha_{\rm diff.} = -1.455}$ and $\mathrm{\alpha_{\rm don-diff.} = -1.468}$. The resulting exponents are slightly smaller than previously reported values of $\mathrm{\approx - 5/3}$ \cite{SORN_Paper}. 
%%% plot_syn_lifetimes.py
\includegraphics[width=\textwidth]{../../plots/syn_topology/syn_lifetimes_new.png}
\caption{Distribution of lifetimes of recurrent excitatory synaptic connections. Linear fits ignoring falloff (only using bins with $\rm t \leq 150 \, s$) for large synaptic lifetimes resulted in a power-law exponent of $\mathrm{\alpha_{\rm diff.} = -1.455}$ (standard error $\mathrm{\sigma_{\rm diff.} = 5.52 \cdot 10^{-4}}$) for diffusive homeostasis and $\mathrm{\alpha_{\rm don-diff.} = -1.468}$ (standard error $\mathrm{\sigma_{\rm don-diff.} = 5.70 \cdot 10^{-4}}$) for non-diffusive homeostasis.}
\label{Syn_Lifetimes}
To get an approximative explanation for the observed slope in the region of small lifetimes, we note that short synaptic lifetimes are most likely coming from synapses whose weights did not reach large values compared to the initial weight. The effect of multiplicative normalization is small compared to additive STDP in this regime of small weights. Small weights are therefore determined by additive quasi-random (due to their weak coupling to the postsynaptic neuron) fluctuations, which, as a very simple approximation, can be described as a Brownian motion with a certain standard deviation $\mathrm{\sigma}$. The first-hitting time of this Brownian motion is described by a Lévy distribution \cite{Roy_1968}:
p(t,w_0,w_{\rm prune}) = \frac{ | w_0-w_{\rm prune}|}{\sqrt{2\pi \sigma t^3}} \mathrm{exp} \left( \frac{ \left( w_0-w_{\rm prune} \right) ^2}{2\sigma t}\right)
\label{Levy_Dist}
The term $\mathrm{t^{-3/2}}$ dominates for larger $\mathrm{t}$, which approximately resembles the power-law exponents found in Figure \ref{Syn_Lifetimes}. Zheng et al. found a similar slope in the binary version of the SORN and also argued that this could be referred to random-walk behavior \cite{Pengsheng_2013}. Still, understanding the exponent of this distribution is subject to ongoing research and it has been argued that a more precise explanation could be used to relate certain parameters of synaptic plasticity to the value of this exponent. Especially, the balance between long-term potentiation and depression is thought to influence this exponent, where potentiation/depression dominated plasticity leads to a smaller/larger absolute value of the exponent of the distribution.

@@

\begin{figure}
%%% plot_sum_outgoing_weights_dist.py
\includegraphics[width=\textwidth]{../../plots/syn_topology/outgoing_weights_distribution.png}
\caption{Mean of outgoing weights of excitatory neurons in order sorted by magnitude. Dashed lines are means obtained by randomly shuffling synaptic weights. Data was taken from a single weight matrix each, at $\mathrm{t = 1500\,s}$.}
\label{Out_Weight_Mean_Quantile}
\end{figure}

@@

Figure \ref{dynamics_rate_threshold_dyn_mean_field_sweeney} depicts the resulting dynamics of excitatory rates and thresholds for $\mathrm{\alpha=\lbrace 0.5,0.9,1.0\rbrace}$. We ran simulations with other values of $\rm \alpha$ and found that all rates approached the same target rate of $\mathrm{3\,Hz}$ and thresholds moved towards the same configuration for either choice of $\mathrm{\alpha}$ except for the limiting case of $\mathrm{\alpha=1}$. A significant difference only existed within the transient dynamics leading to the steady state. Roughly speaking, a smaller value of $\mathrm{\alpha}$ led to a faster relaxation.
\begin{figure}
%%% plot_sim_mean_field_sweeney.py
\includegraphics[width=\textwidth]{../../plots/rate_threshold_simpl_hip_alpha/rate_threshold_simpl_hip_alpha_comb}
\caption{Dynamics of rates (A) and thresholds (B) of excitatory population of 400 Neurons and different values of $\mathrm{\alpha}$ (see Equations \eqref{dyn_rate_equation} and \eqref{diff_hom_simpl_sweeney}).}
\label{dynamics_rate_threshold_dyn_mean_field_sweeney}
\end{figure}
These findings led us to the conclusion that---apart from the special case $\mathrm{\alpha=1}$, which is the equivalent to instantaneous diffusion---a mean-field model describing the diffusive signal as a mixture of individual activity and all neurons is not suitable for explaining the existence of stable heterogeneous distributions of rates for a broad range of diffusion constants. We did not intent to debase the validity of results of the mean-field model of Sweeney et al. However, we could show that the steady-state solution of this simplified model will result in the same firing rate $\phi_0$ for all neurons, given any value of $\mathrm{\alpha}$ except $\mathrm{\alpha=1}$. This can be seen by setting the left hand side of Equation \eqref{diff_hom_simpl_sweeney} to $0$ (which is necessarily the case in the steady state) and rearranging the equation:
\begin{equation}
(\alpha-1)\frac{\phi_i - \phi_0}{\phi_i} = \alpha \frac{\langle \phi \rangle - \phi_0}{\langle \phi \rangle}
\label{diff_hom_simpl_sweeney_2}
\end{equation}
The right term of the equation is the same for all neurons. Since the left term is monotonically increasing as a function of $\phi_i$ as long as $\mathrm{\alpha<1}$, only one specific solution $\phi_i = \Phi$ for all $i$ exists that equals the given term on the right. This implies $\langle \phi \rangle = \Phi$. Thus,
\begin{equation}
(\alpha-1)\frac{\Phi - \phi_0}{\Phi} = \alpha \frac{\Phi - \phi_0}{\Phi}
\label{diff_hom_simpl_sweeney_3}
\end{equation}
which is only fulfilled for $\Phi = \phi_0$.

@@

\section{Discussion}
The investigations and results presented within this thesis should be contextualized within two topics of research: First, the dynamic properties and mechanisms of neural homeostasis and how concepts of control theory should be applied to properly model activity regulation in the brain. And second, if, and if so, how much the particular distribution of firing rates present within a network affects neural plasticity and structure. The focus of this thesis was put onto the first topic. In the introduction, we laid out the conceptual problems of single-cell homeostasis in the context of linear control theory. Though some experimental results suggested that individual homeostatic targets might be a reasonable theoretic approach \cite{OConnor_2010,Mizuseki_2013,Hengen_2016}, we wanted to find a possibility of achieving a broad and heavy-tailed distribution of firing rates and stable network activity without the need to artificially tune individual homeostatic targets to fit the desired distribution. By introducing the notion of diffusive homeostasis, Sweeney et al. provided a theoretical approach that seemed to fulfill this demand. A large part of our research thus focused on the analysis of the dynamics and statistics of network activity that resulted from implementing the mechanism introduced in \cite{Sweeney_Paper}.

The second part of our investigations consisted of checking for possible effects of greater firing-rate heterogeneity onto the network topology, based on findings of earlier research on the SORN. It has been claimed that the mechanisms included in the SORN and LIF-SORN constitute a minimal set of rules required for the emergence of non-random topological features and spine-size dynamics resembling experimental data \cite{SORN_Paper,Pengsheng_2013}. We thus aimed for preserving these features under diffusive homeostasis.

We shall now discuss the presented results in further detail in the following sections.

@@

\subsection{The Firing Rate Distribution} \label{Fir_Rate_Dist_Discussion_Section}
By implementing diffusive homeostasis into the LIF-SORN model we managed to achieve a broader, skewed firing rate distribution while preserving appreciated features of network topology. Though the distribution did approximately resemble a log-normal-like shape, it did not perfectly match our expectations. In order to acquire a deeper understanding of the mechanisms that determine these statistics we derived an analytical expression that allowed us to reliably predict the steady state of firing rates based on the spatial structure of excitatory neurons. We would like to emphasize that this explanation differs from common approaches of understanding the emergence of heavy-tailed statistics of firing rates in recurrent networks. Theoretical studies on the behavior of recurrent networks are largely based on the assumption that the excitability of neurons is not individually fine-tuned but rather randomly distributed and adjusted globally to achieve a desired mean firing rate. Predictions about the resulting activity are then made based on (potentially simplified) statistical features of recurrent connectivity and the specific form of the neuronal transfer function used in the model, see e.g. \cite{Roxin_Firing_Rate_Distribution,Vreeswijk1998,Koulakov_2009}. Our interpretation of our results leads to the conclusion that this canonical approach matches the limiting case of instant diffusion: While Figures \ref{Fir_Rate_Dist_Instant_compare} and \ref{Fir_Rate_Dist_Compare} did not reveal many differences in terms of overall statistics, the decorrelation discussed in Section \ref{Section_Rand_Mat_vs_Sim} and plotted in Figure \ref{Corr_Coeff_vs_D} suggests that increasing the diffusion constant goes along with a transition between a spatially determined configuration of excitatory firing rates and a network topology-determined behavior.

@@

\subsection{Network Topology}

\subsubsection{Robustness of Synaptic Weight- and Lifetime Statistics}
The results of section \ref{Syn_Weight_Dist_Section} and \ref{Section_Syn_Lifetimes} have shown that both features, log-normal distribution of synaptic weights and a power-law distribution of synaptic lifetimes, can be reliably reproduced under our new homeostatic model. Relating this to the main result of Section \ref{Fir_Dist_Section}, it does not appear to be the case that a greater heterogeneity of firing activity has a significant influence onto the overall statistics of neural connectivity (though on a more detailed level, structural differences could be observed, as discussed in the next Section).

The precise spiking activity of neurons does certainly influence weight dynamics due to STDP and learning processes in the brain depend on this relation between neural activity and synaptic structure. However, one might doubt that this non-randomness on a small scale is required to explain overall statistical features of neural connectivity. As mentioned in Section \ref{Syn_Weight_Dist_Section}, theoretical models based on stochastic processes were able to give an explanation for the presence of log-normal weight distributions \cite{Loewenstein_Spine_Sizes,Statman_Synapses_2014}. Individual spikes in this context rather act as initiators of approximately random additive changes of synaptic efficacy. Since diffusive homeostasis kept the population activity at the same value of $\rm 3\, Hz$ as for non-diffusive homeostasis, the amount of additive weight changes due to STDP is the same. Treating STDP-events as random, this equality gives an idea of why no significant statistical differences could be observed for weight distributions and synaptic lifetimes.

On the other hand, persistent differences within firing rates did indeed lead to noticeable topological difference that shall be discussed in the following section

\subsubsection{Strong Differences in Outgoing Weights}
Even though our investigations on network topology were rather intended as a form of sanity check with respect to properties already found in earlier versions, we found a differences compared to non-diffusive homeostasis:
The emergence of significantly above-average mean outgoing weights for a small subset of neurons, see section \ref{Section_Mean_outgoing_Weights}. By comparing the given mean values to statistics acquired by randomly shuffling recurrent excitatory weights and running a statistical test(see Figure \ref{Out_Weight_Mean_Quantile}), we found that this effect is indeed significant compared to chance. Thus, we identify the over-representation of strong outgoing weights as a non-random effect. However, we would like to emphasize that diffusive homeostasis is only indirectly responsible in the sense that it allows for a sustained heterogeneity among firing rates.
